---
title: "a4"
author: "Yuhan"
date: "2019/11/23"
output: 
  pdf_document:
    toc: true
    latex_engine: xelatex
---

\pagebreak

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

# Smoking


## Introduction

Given data from 2014 American National Youth Tobacco Survey. It is known that age at which first cigarette smoking is known to be eariler for males than females, earlier in rural areas than urban areas, and to vary by ethnicity. It is likely that significant variation exists between US states and between schools from the same states.

So we are interested in 2 hypothesis:

1. Geographic variation(between states) in the mean age children first try cigarettes is substantially greater than variation among schools. As a result, tobacco control programs should target the states with the earliest smoking ages and not concern themselves with finding particular schools where smoking is a problem. 

2. First cigarette smoking has a flat hazard function, or in other words is a first order Markov process. This means two non-smoking children have the same probability of trying cigarettes within the next month, irrespective of their ages but provided the known confounders (sex, rural/urban, etnicity) and random effects (school and state) are identical.

## Methods

Since one can only experience first cigarette smoking once, Weibull distribution is chosen for this study. Model is given below.

$Y_{ijk}\sim Weibull(\rho_{ijk,k})$

$\rho_{ijk} = exp(-\eta_{ijk})$

$\eta_{ijk}=X_{ijk}\beta+U_i+V_{ij}$

$U_i\sim N(0,\sigma_U^2)$

$V_i\sim N(0,\sigma_V^2)$

Response variable $Y_{ijk}$ represents the age when a person first try smoking, where $i$ stands for states, $j$ stands for school, $k$ stands for individual.

Event = 0 stands for right censoring, means the given event haven't been observed yet. In this case, the event is that the participants first try smoking. 

Event = 1 stands for we observed the event.

Event = 2 stands for left censoring, which means we didn't observe the event because the event happened before we start to observe.

$U_i$ and $V_{ij}$ are random effects for states and school respectively.

Besiedes, we have ethnicity, gender and residence area(rural or urban) as fixed effect, which is represented by $X_{ijk}\beta$ in the model.

There are 3 prior we need to choose

1. For shape parameter of weibull distribution, which is $\alpha$

2. For $U_i$, we use pc prior.

3. For $V_{ij}$, we use pc prior.

```{r}
load('smoke.RData')
smoke = smoke[smoke$Age > 9, ]
forInla = smoke[, c("Age", "Age_first_tried_cigt_smkg", "Sex", "Race", "state", "school", "RuralUrban")]
forInla = na.omit(forInla)

forInla$school = factor(forInla$school)

library("INLA")
forSurv = data.frame(time = (pmin(forInla$Age_first_tried_cigt_smkg, forInla$Age) - 4)/10, event = forInla$Age_first_tried_cigt_smkg <= forInla$Age)
# left censoring

forSurv[forInla$Age_first_tried_cigt_smkg == 8, "event"] = 2

smokeResponse = inla.surv(forSurv$time, forSurv$event)

fitS2 = inla(smokeResponse ~ RuralUrban + Sex + Race +
               f(school, model = "iid", hyper = list(prec = list(prior = "pc.prec", param = c(log(1.5), 0.05))))+
               f(state, model = "iid", hyper = list(prec = list(prior = "pc.prec", param = c(1, 0.05)))),
             control.family = list(variant = 1, hyper = list(alpha = list(prior = "normal", param = c(log(1),(2/3)^(-2))))),
             control.mode = list(theta = c(8, 2, 5), restart = TRUE),
             data = forInla, family = "weibullsurv", verbose = TRUE)

fitS2$mode$theta

knitr::kable(rbind(fitS2$summary.fixed[, c("mean","0.025quant","0.975quant")], Pmisc::priorPostSd(fitS2$summary[, c("mean", "0.025quant", "0.975quant")])))

knitr::kable(rbind(fitS2$summary.fixed[, c("mean","0.025quant","0.975quant")], Pmisc::priorPostSd(fitS2$summary[, c("mean", "0.025quant", "0.975quant")])))

fitS2$priorPost = Pmisc::priorPost(fitS2)
for (Dparam in fitS2$priorPost$parameters) {
  do.call(matplot, fitS2$priorPost[[Dparam]]$matplot)
  }


xSeq = seq(5,100,len=1000)
cumHazC = (xSeq / exp(fitS2$coef['(Intercept)']))^(1/fitS2$scale)

plot(xSeq, cumHazC, col='blue',type='l', log='y', ylim=c(0.01, 5), xlab='years', ylab = 'cum haz', xlim = c(25,95))
hazEst = survfit(Surv(life, ones) ~ 1, data=dat)
lines(hazEst, fun='cumhaz')
```

## Result

Both hypothesis are false. 


\pagebreak

# Death on the roads


## Introduction

Given the data about all of the road traffic accidents in the UK from 1979 to 2015. We want to invesitgate whether women tend to be, on average, safer as pedestrians than men, particularly as teenagers and in early adulthood. The data consist of all pedestrians involved in motor vehicle accidents with either fatal or slight injuries except pedestrians with moderate injuries.

## Methods

$pr(Y_i=1|X_i)=\lambda_i$

$\log[\lambda_i/(1-\lambda_i)]=\beta_0+\sum_{p=1}^{P}X_{ip}\beta_p$



## Results



```{r}
library("survival")
#theClogit = clogit(y ~ age + age:sex + strata(strata), data = x, )

pedestrians = readRDS('pedestrians.rds')
pedestrians = pedestrians[!is.na(pedestrians$time),]
pedestrians$y = pedestrians$Casualty_Severity == "Fatal"
pedestrians$timeCat = format(pedestrians$time, "%Y_%b_%a_h%H")
pedestrians$strata = paste(pedestrians$Light_Conditions, pedestrians$Weather_Conditions, pedestrians$timeCat)

theTable = table(pedestrians$strata, pedestrians$y)
onlyOne = rownames(theTable)[which(theTable[, 1] == 0 | theTable[, 2] == 0)]
x = pedestrians[!pedestrians$strata %in% onlyOne, ]

theCoef = rbind(as.data.frame(summary(theClogit)$coef), `age 26 - 35` = c(0, 1, 0, NA, NA))
theCoef$sex = c("Male", "Female")[1 + grepl("Female", rownames(theCoef))]
theCoef$age = as.numeric(gsub("age|Over| - [[:digit:]].*|[:].*", "", rownames(theCoef)))
theCoef = theCoef[order(theCoef$sex, theCoef$age), ]
matplot(theCoef[theCoef$sex == "Male", "age"], exp(as.matrix(theCoef[theCoef$sex == "Male", c("coef", "se(coef)")]) %*% Pmisc::ciMat(0.99)), log = "y", type = "l", col = "black", lty = c(1, 2, 2), xaxs = "i", yaxs = "i")
matplot(theCoef[theCoef$sex == "Female", "age"], exp(as.matrix(theCoef[theCoef$sex == "Female", c("coef", "se(coef)")]) %*% Pmisc::ciMat(0.99)), log = "y", type = "l", col = "black", lty = c(1, 2, 2), xaxs = "i")
```


