---
title: "A1"
author: "Yuhan"
date: "2019/9/29"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dtt)
boats <- matrix(scan('boats.txt'),ncol=256,byrow=T)
```

## Q1
### a.
${D_m}^{-1}{A_m}^T\hat{Z}A_n{D_n}^{-1}={D_m}^{-1}{A_m}^TA_mZ{A_n}^TA_n{D_n}^{-1}$

since ${A_n}^TA_n=D_n$

the above notation$={D_m}^{-1}D_mZD_n{D_n}^{-1}$

$D_n$ is diagonal given

${D_m}^{-1}D_m=D_n{D_n}^{-1}=I$

so ${D_m}^{-1}{A_m}^T\hat{Z}A_n{D_n}^{-1}={D_m}^{-1}D_mZD_n{D_n}^{-1}=IZI=Z$

### b.

```{r, include=TRUE}
###########Given function,labelled as denoise1()
denoise1 <- function(dctmat,lambda) {
  # if lambda is missing, set it to the 0.8 quantile of abs(dctmat)
  if(missing(lambda)) lambda <- quantile(abs(dct),0.8)
  # hard-thresholding
  a <- dctmat[1,1]
  dctmat1 <- ifelse(abs(dctmat)>lambda,dctmat,0)
  dctmat1[1,1] <- a
  # inverse DCT to obtain denoised image "clean"
  clean <- mvdct(dctmat1,inverted=T)
  clean <- ifelse(clean<0,0,clean)
  clean <- ifelse(clean>1,1,clean)
  clean
}

##########Modified Function,,labelled as denoise2()
denoise2 <- function(dctmat,lambda) {
  # if lambda is missing, set it to the 0.8 quantile of abs(dctmat)
  if(missing(lambda)) lambda <- quantile(abs(dct),0.8)
  # soft-thresholding
  a <- dctmat[1,1]
  dctmat1 <- sign(dctmat)*pmax(abs(dctmat)-lambda,0)
  dctmat1[1,1] <- a
  # inverse DCT to obtain denoised image "clean"
  clean <- mvdct(dctmat1,inverted=T)
  clean <- ifelse(clean<0,0,clean)
  clean <- ifelse(clean>1,1,clean)
  clean
}
```




### c.
```{r, echo = FALSE}
mvdct.boats <- mvdct(boats)
boats10 <- denoise1(mvdct.boats, lambda = 10)
boats20 <- denoise1(mvdct.boats, lambda = 20)
boats30 <- denoise1(mvdct.boats, lambda = 30)
boats40 <- denoise1(mvdct.boats, lambda = 40)

image(boats10,axes = F,col = grey(seq(0,1,length=256)),sub='lambda = 10',asp=1)
image(boats20,axes = F,col = grey(seq(0,1,length=256)),sub='lambda = 20',asp=1)
image(boats30,axes = F,col = grey(seq(0,1,length=256)),sub='lambda = 30',asp=1)
image(boats40,axes = F,col = grey(seq(0,1,length=256)),sub='lambda = 40',asp=1)
```

```{r,echo = FALSE}
mvdct.boats <- mvdct(boats)
boats10 <- denoise2(mvdct.boats, lambda = 10)
boats20 <- denoise2(mvdct.boats, lambda = 20)
boats30 <- denoise2(mvdct.boats, lambda = 30)
boats40 <- denoise2(mvdct.boats, lambda = 40)

image(boats10,axes = F,col = grey(seq(0,1,length=256)),sub='lambda = 10',asp=1)
image(boats20,axes = F,col = grey(seq(0,1,length=256)),sub='lambda = 20',asp=1)
image(boats30,axes = F,col = grey(seq(0,1,length=256)),sub='lambda = 30',asp=1)
image(boats40,axes = F,col = grey(seq(0,1,length=256)),sub='lambda = 40',asp=1)
```

For both method, as value of lamba increases, contrast of picture decreases. In addition, contrast of hard-threshold approach looks better to me.



## 2

### a.
$E_X(s^X) = E_{U+2V}(s^{U+2V})=E_U(s^UE_{2V}(s^{2V}))=E_U(s^X)*E_{2V}(s^{2V})$

$E(s^U)=E(s^{Poisson(\lambda_u)})= exp(\lambda_u(s-1))$

$E(s^{2V}) = exp(\lambda_v(s^2-1))$

$E_X(s^X) =E_U(s^X)*E_{2V}(s^{2V})=exp(\lambda_u(s-1))exp(\lambda_v(s^2-1))=exp(\lambda_u(s-1)+\lambda_v(s^2-1))$

### b.



###c.




